\documentclass[runningheads,a4paper]{llncs}
\usepackage{makeidx}
\usepackage{amssymb}
\usepackage{amscd, amsfonts}
\usepackage{verbatim}
\usepackage{listings}
\usepackage{commath}
\usepackage{float}
\usepackage{tikz}
\usepackage{stmaryrd}

\usepackage{lmodern}
\usepackage[T1]{fontenc}

\newcommand{\dL}{\mathsf{d}\mathcal{L}}

\lstset{
  mathescape,
  columns=fullflexible,
  basicstyle=\small\fontfamily{lmvtt}\selectfont
}

% Allow display breaks
\allowdisplaybreaks[1]

\begin{document}

\author{}
\institute{}

\title{ARCH-COMP19 Category Report: Hybrid Systems Theorem Proving}

\maketitle

\paragraph{Description of the system.}

HHL prover~\cite{WZZ15} is an interactive theorem prover for verifying
hybrid systems modelled by Hybrid CSP (HCSP)~\cite{He94,ZWR96}. It
implements the Hybrid Hoare Logic (HHL)~\cite{LLQZ10}, a Hoare style
logic for reasoning about HCSP, in the proof assistant
Isabelle/HOL~\cite{isabelle}.

HCSP extends CSP by introducing continuous variables, differential
equations, and interruptions by domain boundary and
communication. Given an HCSP process $P$, an HHL specification takes
the form $\{\mathit{Pre}\}~P~\{\mathit{Post}; \mathit{HF}\}$, where
$\mathit{Pre}$ and $\mathit{Post}$ are pre-/post-conditions in
first-order logic, and $\mathit{HF}$ is a history formula in duration
calculus~\cite{ZH04} for specifying time-related properties held
throughout the whole execution interval. HHL defines a set of proof
rules for deducing such specifications for HCSP. HHL prover formalizes
HCSP and HHL for proving partial correctness of hybrid systems.

For reasoning about differential equations, HHL includes a proof rule
that reduces the specification of the continuous evolution to be
proved to the synthesis problem of differential invariants for the
corresponding differential equations. HHL prover resorts to an
external invariant generator based on quantifier elimination or
sum-of-squares (SOS) relaxation, to automatically solve the unproven
constraints containing unknown differential invariants. The invariant
generator relies on the solvers for quantifier elimination and
semi-definite programming for constructing differential invariants. In
the newest version of HHL prover, we removed the dependency on
Mathematica for the SOS-based invariant generator.

\paragraph{Ghost rule and cut rule.}

For the newest version of HHL prover, we also borrow some idea
from~\cite{Platzer18} and add the differential cut and ghost rules to
the HHL framework. Differential cut rule strengthens the domain of
differential equations by an invariant property proved to hold
throughout the continuous evolution, while the differential ghost rule
adds new continuous variables with new differential equations without
affecting the original differential equations. These rules make
reasoning about differential equations more flexible. For differential
invariants, the current version of HHL prover resorts to external
tools for both invariant verification and generation. Stating proof
rules for reasoning about differential invariants within Isabelle is
left open for future work.

\paragraph{Verification of differential invariants.}

For the newest version of the HHL Prover, we added an invariant
checker using quantifier elimination from the external tool
Redlog. Invoking Redlog from Isabelle follows the same pattern as for
invariant generation. First, the goal to be proved, which may contain
both entailments (of the form $p\implies q$) and preservation of
invariant by a differential equation (of the form
$\mathtt{exeFlow}(\mathbf{x'}=f(\mathbf{x}),I)\implies I$), is
translated from Isabelle's abstract syntax tree to JSON format. Next,
the JSON file is translated to the input for Redlog using a Python
script. In this step, preservation of invariant is converted to
appropriate entailments involving the domain of evolution and the Lie
derivative of the invariant. Finally, Redlog is invoked on the output
of the Python script and checks each of the resulting
entailments. This sequence is implemented as a bash script and invoked
by an oracle in Isabelle, which checks the final output produced by
Redlog. The entire process is automatic, after the user supplies the
invariant in usual mathematical notation, and a list of constants that
are relevant for quantifier elimination.

\paragraph{Example: design shapes and nonlinear continuous models.}

We translated some of the examples in the first two parts of the
benchmarks to HCSP. There are some essential differences between the
semantics of $\dL$ and HCSP. In particular, in HCSP, evolution by a
differential equation cannot stop before reaching the boundary (or
interrupted by a communication). Also, there is no assignment to an
arbitrary value ($x := *$) in HCSP. As a result, some of the examples
cannot be translated naturally. For those examples that can be
translated naturally, we proved some of them in Isabelle, with the
help of invariant checking using Redlog (details to be completed by
Mar 24).

\paragraph{Example: Roller coaster.}

We converted the roller coaster example \cite{coasterx} to HCSP. The
conversion is natural, as the differences between $\dL$ and HCSP does
not produce any problems. The proof makes use of invariant checking
using Redlog, as well as the newly added differential ghost rule. The
entire Isabelle theory (including the model, specification, and proof
for all ten parts of the example) is 1141 lines long.

\paragraph{Example: lunar lander control program.}

The lunar lander control program is a closed loop system, which is
composed of the lander's dynamics and the guidance program for the
slow descent phase. The guidance program is executed periodically with
a fixed sampling period. At each sampling point, the current state of
the lander is measured by inertial measurement unit or various
sensors. Processed measurements are then input into the guidance
program, which outputs control commands, e.g. the magnitude and
direction of thrust, to be imposed on the lander's dynamics in the
following sampling cycle.

The mathematical description of the lander's dynamics as well as the
guidance program of the slow descent phase can be found
in~\cite{ZYZG14,ZhanWZ16}. The entire Isabelle theory including the
model, specification, and proof for the entire example is 327 lines
long. By applying HHL prover, the unproven subgoals related to
differential invariants are transformed to a set of SOS constraints
with respect to the user-defined invariant template, and then the
SOS-based invariant generator is invoked on these constraints to
synthesize a satisfying invariant.


\bibliographystyle{splncs04}
\bibliography{main}

\end{document}
